\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-07-13 coq-example.tex
Adding comments.

\begin{wrapfigure}{r}{0.5\textwidth}
\begin{lstlisting}[language=Coq,frame = single, escapeinside={@}{@}]
Family STLC_prod.
 Extend FInductive ty : Set :=
  | ty_prod : ty -> ty -> ty.
 Extend FInductive tm : Set :=
  | tm_prod : tm -> tm -> tm  
  | tm_pi1 : tm -> tm  
  | tm_pi2 : tm -> tm.
 Extend Family subst_internal.
  Final Field tm_prod := ..
  Final Field tm_pi1 := ..
  Final Field tm_pi2 := ..
 EndFamily.

@@

\subsection{Abstract Interpreters for \texttt{Imp}}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, here, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property.
Then we extend the base language with new features, and we instantiate
the postulation on computation for
abstract interpreters.
Thanks to the compilation to Coq module, we can directly run the resulting
abstract interpreters.

>>>>

\begin{wrapfigure}{r}{0.5\textwidth}\YZ{It is more convincing if you have four or five family mixin together, positive product, negative product, sum type, boolean, and natural number, general recursion and one more}
\begin{lstlisting}[language=Coq,frame = single, escapeinside={@}{@}]
Family STLC_prod.
 Extend FInductive ty : Set :=
  | ty_prod : ty -> ty -> ty.
 Extend FInductive tm : Set :=
  | tm_prod : tm -> tm -> tm  
  | tm_pi1 : tm -> tm  
  | tm_pi2 : tm -> tm.
 Extend Family subst_internal.
  Final Field tm_prod := ..
  Final Field tm_pi1 := ..
  Final Field tm_pi2 := ..
 EndFamily.

@@

\subsection{Abstract Interpreters for \texttt{Imp}}\YZ{Try to have two runnable interpreter -- choose another simple one like constant propagation. and remove LangMore}
The second example is adapted and modified from \citet{zm2017}.
Contrary to our first example, here, we use a big-step interpreter with fuel
as the operational semantics of an imperative language with
side effects, and we specify the abstract interpretation and prove its
soundness, with some postulation on both computation and property.
Then we extend the base language with new features, and we instantiate
the postulation on computation for
abstract interpreters.
Thanks to the compilation to Coq module, we can directly run the resulting
abstract interpreters.

### 2022-07-13 coqimpl.tex
Adding comments.

Once the user closes a family with "EndFamily", they can access its fields by using module syntax. For example, they can access "A.a" to get "1" after "Family A" is defined. This is achieved by a complete compilation from a family to a module, via aggregating the list of the modules in the internal representation: the plugin will repeatedly using Coq's "Include" command as the "Module A" in \cref{fig:plugin-example1} illustrates. Now, the user can use "Module A" anywhere as they are using "Family A".
\YZ{The reader wonders why this aggregation is needed. 
Seems like the place to talk about how the programmer can use A outside Family A.}\EDJreply{Please check.}

@@

\begin{figure}[!htb]
  \begin{minipage}[t]{0.32\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family B.
  FInductive b : Set 
    := tt : b | ff : b.
  Family neg_handler. 
    Field tt : self_B.b 
      := self_B.ff.
    Field ff : self_B.b 
      := self_B.tt.
  EndFamily.
  FRecursor neg 
    about self_B.b 
    motive (fun _ => self_B.b)
    using self_B.neg_handler
    by _rec.
  Field example := 
    self_B.neg self_B.tt. 
EndFamily.
Family B2 extends B.
  Extend FInductive b : Set 
    := uu : b.
  Extend Family neg_handler.
    Field uu : self_B2.b 
      := self_B2.uu.
  EndFamily. 
EndFamily.
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{0.65\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
  (* Abstraction for Inductive Type *)
Module Type b_3 (self_B: EmptySig_4).
Parameter (b : Set). Parameter (tt ff : b). End b_3.      
Module Type b_6 := b_3 EmptyMod.
  (* Recursor to Set for b *)
Module b_rec_12 (self_B: b_6).
Definition __recursor_type_b_rec :=
  forall P : self_B.b -> Set,
  P self_B.tt -> P self_B.ff 
  -> forall __i : self_B.b, P __i. End b_rec_12. 
  (* Field B.neg_handler.ff *)
Module ff_24 (self_B: b_6) 
             (self_neg_handler: tt_23 self_B).
Definition ff : self_B.b := self_B.tt. End ff_24.
  (* Intermediate Module solely for type checking *)
Module v_33_34 (self_B: neg_handler_26).
Include b_rec_12 self_B.
Parameter (recursor_for_type_checking : __recursor_type_b_rec).
Definition term_for_type_checking :=
  recursor_for_type_checking (fun _ : self_B.b => self_B.b)
    self_B.neg_handler.tt
    self_B.neg_handler.ff.  
End v_33_34.
\end{minted}
  \end{minipage}
\caption{Example Code for Inductive Type}\label{fig:plugin-example2}
\end{figure}

@@

\textbf{Hack: Global Reasoning.}
% explain the places using ``Closing Fact''
% 1. No need to export partial recursor
% 2. Computational rule for the total recursor
Unfortunately, there are cases an extensible proof brings more boilerplate code. For example, in STLC, we expect
``\mintinline{Coq}{~ value (tm_app x y)}'' to hold in all future
extensions of STLC---applications will never be considered as values
by any extension.
We know the correct way to prove the proposition
\mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use
"FRecursor" to inductively reason "value (tm_app x y)". However, this
means that every time we extend "value" with new value forms, we will
need to extend corresponding inductive cases for "app_not_value". We can
imagine the newly extended proof to be boring case analysis that the
newly added value forms are not "tm_app". What's worse, similar
scenario can happen on other statements like \mintinline{Coq}{~ value
(tm_if cond x y)}. Everything here is extended correctly but we end up
having boilerplate code, \textbf{semantically}.
\begin{figure}[!htb]
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC.
  FInductive tm : Set := ... 
  FInductive value : tm -> Prop := ... 
  Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
      by { intros x y H; inversion H; eauto }.
  Fail Ltac inv := 
    match goal with 
    | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
    (* Fail: value, tm_app, value_not_app unfound *)
    end. 
  MetaData tactic1.
    Ltac inv := (* ... the same definition as above ... *)
  EndMetaData.
EndFamily.
\end{lstlisting}  
\caption{Example for Global Reasoning and MetaData}\label{fig:plugin-example-global-reasoning-meta-data}
\end{figure}

>>>>

Once the user closes a family with "EndFamily", they can access its fields by using module syntax. For example, they can access "A.a" to get "1" after "Family A" is defined.\YZ{1. move some of sec 3.2 ~> 3.1; 2. hint vaguely on how to do type checking in sec 3.1 (user interface).. metatheory part should also mention this part in sec 3.1} This is achieved by a complete compilation from a family to a module, via aggregating the list of the modules in the internal representation: the plugin will repeatedly using Coq's "Include" command as the "Module A" in \cref{fig:plugin-example1} illustrates. Now, the user can use "Module A" anywhere as they are using "Family A".
\YZ{The reader wonders why this aggregation is needed. 
Seems like the place to talk about how the programmer can use A outside Family A.}\EDJreply{Please check.}

@@

\begin{figure}[!htb]\YZ{Change this one into STLC example, to see the layout}
  \begin{minipage}[t]{0.32\linewidth}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family B.
  FInductive b : Set 
    := tt : b | ff : b.
  Family neg_handler. 
    Field tt : self_B.b 
      := self_B.ff.
    Field ff : self_B.b 
      := self_B.tt.
  EndFamily.
  FRecursor neg 
    about self_B.b 
    motive (fun _ => self_B.b)
    using self_B.neg_handler
    by _rec.
  Field example := 
    self_B.neg self_B.tt. 
EndFamily.
Family B2 extends B.
  Extend FInductive b : Set 
    := uu : b.
  Extend Family neg_handler.
    Field uu : self_B2.b 
      := self_B2.uu.
  EndFamily. 
EndFamily.
\end{lstlisting}
  \end{minipage}
  \begin{minipage}[t]{0.65\linewidth}
\begin{minted}[fontsize=\footnotesize,escapeinside=@@]{Coq}
  (* Abstraction for Inductive Type *)
Module Type b_3 (self_B: EmptySig_4).
Parameter (b : Set). Parameter (tt ff : b). End b_3.      
Module Type b_6 := b_3 EmptyMod.
  (* Recursor to Set for b *)
Module b_rec_12 (self_B: b_6).
Definition __recursor_type_b_rec :=
  forall P : self_B.b -> Set,
  P self_B.tt -> P self_B.ff 
  -> forall __i : self_B.b, P __i. End b_rec_12. 
  (* Field B.neg_handler.ff *)
Module ff_24 (self_B: b_6) 
             (self_neg_handler: tt_23 self_B).
Definition ff : self_B.b := self_B.tt. End ff_24.
  (* Intermediate Module solely for type checking *)
Module v_33_34 (self_B: neg_handler_26).
Include b_rec_12 self_B.
Parameter (recursor_for_type_checking : __recursor_type_b_rec).
Definition term_for_type_checking :=
  recursor_for_type_checking (fun _ : self_B.b => self_B.b)
    self_B.neg_handler.tt
    self_B.neg_handler.ff.  
End v_33_34.
\end{minted}
  \end{minipage}
\caption{Example Code for Inductive Type}\label{fig:plugin-example2}
\end{figure}

@@

\textbf{Hack: Global Reasoning.}\YZ{This is more like an escape hatch. Escape from family poly into vanilla Coq. Maybe these two hacks should be grouped into one section}
% explain the places using ``Closing Fact''
% 1. No need to export partial recursor
% 2. Computational rule for the total recursor
Unfortunately, there are cases an extensible proof brings more boilerplate code. For example, in STLC, we expect
``\mintinline{Coq}{~ value (tm_app x y)}'' to hold in all future
extensions of STLC---applications will never be considered as values
by any extension.
We know the correct way to prove the proposition
\mintinline{Coq}{value_not_app : value (tm_app x y) -> False.} is to use
"FRecursor" to inductively reason "value (tm_app x y)". However, this
means that every time we extend "value" with new value forms, we will
need to extend corresponding inductive cases for "app_not_value". We can
imagine the newly extended proof to be boring case analysis that the
newly added value forms are not "tm_app". What's worse, similar
scenario can happen on other statements like \mintinline{Coq}{~ value
(tm_if cond x y)}. Everything here is extended correctly but we end up
having boilerplate code, \textbf{semantically}.
\begin{figure}[!htb]\YZ{Make this part concise. (Just say this is a continuation of the earlier STLC example)}
\begin{lstlisting}[language=Coq,  escapeinside={@}{@}]
Family STLC.
  FInductive tm : Set := ... 
  FInductive value : tm -> Prop := ... 
  Closing Fact value_not_app : forall x y, ~ value (tm_app x y) 
      by { intros x y H; inversion H; eauto }.
  Fail Ltac inv := 
    match goal with 
    | [h : value (tm_app _ _) |- ] => destruct (value_not_app _ _ h)
    (* Fail: value, tm_app, value_not_app unfound *)
    end. 
  MetaData tactic1.
    Ltac inv := (* ... the same definition as above ... *)
  EndMetaData.
EndFamily.
\end{lstlisting}  
\caption{Example for Global Reasoning and MetaData}\label{fig:plugin-example-global-reasoning-meta-data}
\end{figure}
